---
title: "Initial sticky card data cleaning"
author: "Caroline + Ariella"
date: "10/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

# Load packages
```{r}
library(tidyverse)
```

# Analysis plan

## what questions do we want to answer?
For each of these things...

timing
diversity
number (abundance)
biomass
dispersion onto land

we can ask:
what is X?
how does X differ with fish presence?
how does X differ with water temperature?

## what do we need before we can answer them?
- insect counts + ids for each trap
- trap information (date, distance from shore, direction, lake, etc)
- lake information (water temp, fish status, etc)

### Insect counts and IDs
- load raw insect data
- make sure columns are in the right format
- check notes/other columns



# Load and inspect data

```{r}

lab_stickycards = readxl::read_excel("data/raw_data/stickycard_counted_inlab.xlsx")

str(lab_stickycards)
dim(lab_stickycards)

# add columns to split info up
lab_data = lab_stickycards %>% 
  mutate(cole_weevil = NA,
         cole_flower = NA,
         .after = Coleoptera) %>% 
  mutate(sim_sm = NA,
         sim_md = NA, 
         sim_lg = NA,
         .after = Simuulidae) %>% 
  mutate(terrestrial = NA) %>% 
  mutate(index = row_number(), .before = Datasheet) #add a row index for each observation

#write new file
#write_csv(lab_data, "data/cleaned_data/drafts/lab_stickycards.csv")

```

# Clean data
Next step: open lab_stickycards.csv in Excel
  - save as a new version w your initials and date
  - for each row, check notes + other columns and move all info from there to appropriate other columns
  - see the following rules for how to sort the information - add any more rules that you use below
  - fill in the row numbers that you're done with here
 
ROWS EDITED:

## Ariella: 1-198. Need to check cards: (finished)


Get information about cards that need checking
```{r}
library(tidyverse)

# Make a list of card indexes

cards <- c(27, 30, 31, 32, 35, 36, 41, 42, 44, 45, 85)
df <- read_csv("data/cleaned_data/drafts/lab_stickycards_AK.csv")


# Use list to filter the data frame

card_info <- df %>% 
  filter(index%in%cards)

# Use information to find cards and check them

View(card_info)

```

## Caroline: 199 -end. 
need to check cards for:

```{r}
test = read_csv("data/cleaned_data/drafts/lab_stickycards_CO.csv") %>% 
  filter(index > 198) %>% 
  select(-c(2:11)) %>% 
  filter(is.na(cole_weevil) | is.na(sim_sm) | is.na(terrestrial) | is.na(`Dip-other`))
  
check_cards = read_csv("data/cleaned_data/drafts/lab_stickycards_CO.csv") %>%
  filter(index %in% test$index)
```

## Combine

1. Combine rows 1-198 (Ariella) and 199-end(Caroline) into new file
lab_stickycards_AKCO.csv
2. Rearrange columns to match
3. Use this new data sheet to finish checking cards

```{r}
test = read_csv("data/cleaned_data/lab_stickycards_AKCO.csv") %>% 
  select(-c(2:11)) %>% 
  filter(is.na(cole_weevil) | is.na(sim_sm) | is.na(terrestrial) | is.na(`Dip-other`))
  
check_cards = read_csv("data/cleaned_data/lab_stickycards_AKCO.csv") %>%
  filter(index %in% test$index) %>% 
  select(1:7) %>% 
  mutate(Direction = factor(Direction, levels = c("N", "S", "E", "W"))) %>% 
  arrange(`Collection date`, Lake, Direction, Distance)

View(check_cards)
```


# Guidelines for cleaning

-ignore any exuviae (molts) - this is not a separate insect

For things with no aquatic stage:
- hemiptera, leafhoppers, spiders, neuroptera, other predators: "terrestrial" column (remove/ignore)
- moths/butterflies: add to lepidoptera column
- hymenoptera (wasps/bees): add to hym column

For coleoptera (beetles):
-keep total column
-make columns broken down by group (flower beetles or weevils)

For simuulidae:
-keep total column
-make columns broken down by size (small, med, lg)

## Check columns
- Coleoptera column should be sum of cole_weevil and cole_flower
- Simuulidae column should be sum of sim_sm, sim_md, and sim_lg
- Other column should be sum of terrestrial and Dip-other

```{r}

# Load packages
library(tidyverse)

# Read in data
check_columns = read_csv("data/cleaned_data/lab_stickycards_AKCO.csv")

# Make new object
df = check_columns %>% 
  mutate(cole_good = (cole_weevil + cole_flower == Coleoptera))

# Check for false values in cole_good
which(df$cole_good == FALSE)

# Make new object
df2 = check_columns %>% 
  mutate(sim_good = (sim_sm + sim_md + sim_lg == Simuulidae))

# Check for false values in sim_good
which(df2$sim_good == FALSE)

# Make new object
df3 = check_columns %>% 
  mutate(other_good = (`Dip-other` + terrestrial == Other))

# Check for false values in other_good
which(df3$other_good == FALSE)
df3 = df3 %>% filter(other_good == FALSE)
```
# Results: 
## Coleoptera: #39 "cole_flower" should say 3 instead of 2 -- fixed
## Simuulidae: All correct
## Other: All fixed
