---
title: "Data Analysis for Stable Isotopes"
author: "Ariella Katzman"
date: '2023-02-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
library(janitor)
library(lubridate)
library(car)
library(ggeffects)

```

Given that there are 243 kg soil per square meter, how much N (g) is there per square meter?
```{r}

# Calculate N per square meter soil
soil_iso_data4 <- soil_iso_data4 %>%
  mutate(n_g_m2 = pct_n/100 * 243 * 1000)
# % to ratio, kg N/kg soil * kg soil/m2 * g/kg
View(soil_iso_data4)

```


Interactions between N as a function of OM, elevation, fish status
```{r}
# Try out all possible combos of variables
res.lm6 <- lm(n_g_m2 ~ pct_OM * elevation_m * fish_status, data = soil_iso_data4)
summary(res.lm6)
qqPlot(res.lm6)
ggpredict(res.lm6, terms = c("fish_status", "elevation_m"))

res.lm7 <- lm(n_g_m2 ~ pct_OM * (elevation_m + fish_status), data = soil_iso_data4) 
summary(res.lm7)
qqPlot(res.lm7)

res.lm8 <- lm(n_g_m2 ~ pct_OM + elevation_m + fish_status, data = soil_iso_data4) 
summary(res.lm8)
qqPlot(res.lm8)

res.lm9 <- lm(n_g_m2 ~ pct_OM * elevation_m, data = soil_iso_data4) 
summary(res.lm9)
qqPlot(res.lm9)

res.lm10 <- lm(n_g_m2 ~ pct_OM + elevation_m, data = soil_iso_data4) 
summary(res.lm10)
qqPlot(res.lm10)

# Test which model is the best fit
AIC(res.lm6, res.lm7, res.lm8, res.lm9, res.lm10)
# Model of percent OM, elevation, and fish status with interactions is best fit for the data 
```

Interactions between N as a function of OM, lake pair, fish status
```{r}
# Try out all possible combos of variables
res.lm11 <- lm(n_g_m2 ~ pct_OM * lake_pair * fish_status, data = soil_iso_data4)
summary(res.lm11)
qqPlot(res.lm11)

res.lm12 <- lm(n_g_m2 ~ pct_OM * (lake_pair + fish_status), data = soil_iso_data4) 
summary(res.lm12)
qqPlot(res.lm12)

res.lm13 <- lm(n_g_m2 ~ pct_OM + lake_pair + fish_status, data = soil_iso_data4) 
summary(res.lm13)
qqPlot(res.lm13)

res.lm14 <- lm(n_g_m2 ~ pct_OM * lake_pair, data = soil_iso_data4) 
summary(res.lm14)
qqPlot(res.lm14)

res.lm15 <- lm(n_g_m2 ~ pct_OM + lake_pair, data = soil_iso_data4) 
summary(res.lm15)
qqPlot(res.lm15)

# Test which model is the best fit
AIC(res.lm6, res.lm11, res.lm12, res.lm13, res.lm14, res.lm15)
# Model of percent OM, lake pair, and fish status with interactions is best fit for the data 
# Lake pair better than elevation

# Use ggpredict to estimate amount of N in soil at each lake
ggpredict(res.lm11, terms = c("fish_status", "lake_pair"))

```

ggpredict function gives us an estimate for the amount of nitrogen in soil at each lake
next step: calculate extra nitrogen in terms of fish per square meter
Brook g N/fish = 1.853296 g N/fish
Rainbow g N/fish = 2.228152 g N/fish
Humphreys(H1P): brook
LowerDes(H2P): rainbow
Mesa(H3P): rainbow

How many fish would it take to equal the amount of additional N per sq. meter of soil?
```{r}

soil_iso_data4 %>%
  group_by(lake_pair, fish_status) %>% 
    summarise(n_g_m2_mean = mean(n_g_m2, na.rm = TRUE),
            n_g_m2_sd = sd(n_g_m2, na.rm = TRUE))

soil_iso_data4 %>%
  group_by(lake_pair, fish_status) %>% 
    summarise(n_g_m2_mean = mean(n_g_m2, na.rm = TRUE)) %>% 
         pivot_wider(names_from = fish_status, values_from = n_g_m2_mean) %>%
  mutate(diff = present - absent, 
         fish_type = case_when(lake_pair == "H1" ~ "brook", T ~ "rainbow"), 
         N_g_fish = case_when(fish_type == "brook" ~ 1.853296, T ~ 2.228152),
         fish_per_m2 = diff/N_g_fish) 


# Calculate fish/m2 needed in lakes with fish for that difference in soil N
# H1: g N/m2 soil / g N/fish = 794.7171 fish per sq. meter
1472.846 / 1.853296
# H2: g N/m2 soil / g N/fish = -3.138443 fish per sq. meter
-364.0286 / 
# H3: g N/m2 soil / g N/fish = -3.111286 fish per sq. meter
-382.8322 /  
# K1: g N/m2 soil / g N/fish = 

# K2: g N/m2 soil / g N/fish = 

# Find differences in soil N between lakes with/without fish using predicted values
# H1: present - absent = 1472.846 g N/m2
2672.5186 - 1199.6725
# H2: present - absent = -364.0286 g N/m2 ***absent greater than present
1822.5328	- 2186.5614
# H3: present - absent = 184.69 g N/m2
  # 184.69 g N/m2 / 123.0463 g N/fish = 1.50098 fish per sq. meter

```



Next steps:
- Calculating number of fish needed to equal soil N for each lake
- Calculate number of insects needed to equal soil N for each lake
- Writing results in abstract form
- Making figure