---
title: "Data Exploration for Stable Isotopes"
author: "Caroline + Ariella"
date: '2023-1-25'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Want to learn:
- For plants, percent C and N, d13C, and d15N by:
  - Lake
  - Distance
  - Elevation?
  - Fish presence
- For soils:
  - Percent C and N, d13C and d15N by:
    - Fish
    - Elevation
    - Lake
    - % OM
  - Percent N by percent OM

```{r}
# Load packages
library(tidyverse)
library(janitor)
library(lubridate)
library(car)
library(ggeffects)
library(splines)

# Read in data
plant_iso_data <- read_csv("plant_iso_data.csv") %>% 
  janitor::clean_names()

```


PLANTS
Graphs of percent C and percent N in plants based on distance and fish
```{r}

# Make scatterplot of percent C as a function of distance and fish
plant_iso_data %>% 
  ggplot(aes(x = distance, y = pct_c, color = fish)) +
  geom_jitter() +
  ylim(48, 54) +
  geom_smooth(method="lm")  # fish presence decreases percent C, percent C decreases slightly over distance

# Test significance
res.lm <- lm(pct_c ~ distance + fish, data = plant_iso_data)
summary(res.lm) # fish presence significantly decreases percent C in plants

# Make scatterplot of percent N as a function of distance and fish  
plant_iso_data %>% 
  ggplot(aes(x = distance, y = pct_n, color = fish)) +
  geom_jitter() +
  geom_smooth(method="lm") # fish presence increases percent N, percent N increases with distance

# Test significance
res.lm2 <- lm(pct_n ~ distance * fish, data = plant_iso_data)
summary(res.lm2) # fish presence and distance (but not interaction) significantly increase percent N, distance more than fish presence

# percent N likely affected more by distance bc N is more mobile in soil

## Why does fish presence decrease percent C but increase percent N in plants?
# Fish produce ammonia through protein metabolism which is released through the gills -> increased N in water; more N in plants means less C

```

Graphs of d13C and d15N by distance for plants
```{r}
# Make scatterplot of d13C as a function of distance and fish
plant_iso_data %>% 
  ggplot(aes(x = distance, y = d13c, color = fish)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm3 <- lm(d13c ~ distance, data = plant_iso_data)
summary(res.lm3)
qqPlot(res.lm3$residuals) # residuals very close to normal distribution

# Make scatterplot of d15N as a function of distance and fish
plant_iso_data %>% 
  ggplot(aes(x = distance, y = d15n, color = fish)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm4 <- lm(d15n ~ distance + fish, data = plant_iso_data)
summary(res.lm4)

```

Graphs of percent N, d13C and d15N by elevation for plants
```{r}

# Add in lake site data
plant_iso_data2 <- left_join(plant_iso_data, owens_lake_sites, by = c("lake" = "lake_ID"))
View(plant_iso_data2)

# Make scatterplot of percent N by elevation and fish
plant_iso_data2 %>% 
  ggplot(aes(x = elevation_m, y = pct_n, color = fish)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm5 <- lm(pct_n ~ elevation_m + fish, data = plant_iso_data2)
summary(res.lm5)

# Make scatterplot of d13C by elevation and fish
plant_iso_data2 %>% 
  ggplot(aes(x = elevation_m, y = d13c, color = fish)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm6 <- lm(d13c ~ elevation_m + fish, data = plant_iso_data2)
summary(res.lm6)

# Make scatterplot of d15N by elevation and fish
plant_iso_data2 %>% 
  ggplot(aes(x = elevation_m, y = d15n, color = fish)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm7 <- lm(d15n ~ elevation_m * fish, data = plant_iso_data2)
summary(res.lm7)
qqPlot(res.lm7$residuals) # residuals very close to normal distribution

```


SOILS
```{r}
# Import soils data
complete_iso_data <- read_excel("OwensC 220627 CN A003185.xlsx", sheet = "Samples") %>%
  janitor::clean_names()
View(complete_iso_data)

soil_iso_data <- subset(complete_iso_data, type_of_material=='surface soil') %>%
  mutate(sample_id=as.numeric(str_replace(sample_id, "SS", ""))) # %>%
  # filter(!is.na(sample_id))
View(soil_iso_data)

soil_metadata = readxl::read_xlsx("C:/Users/Ariella Katzman/Downloads/soil_edited_2021.xlsx", 
    sheet=2) %>%
  janitor::clean_names() %>%
  select(sample_ID = sample_description_number_1,
         pct_OM = organic_matter_percent,
         lake, distance_m, veg1_amt:notes) %>%
  mutate(sample_ID = as.numeric(sample_ID))


# Add soil metadata to isotope data
soil_iso_data2 <- left_join(soil_iso_data, soil_metadata, by = c("sample_id" = "sample_ID"))
View(soil_iso_data2)

# Add lake site data to isotope data
soil_iso_data3 <- left_join(soil_iso_data2, owens_lake_sites, by = c("lake" = "lake_ID"))
View(soil_iso_data3)

# Edit column names
colnames(soil_iso_data3)[3] = "total_c_ug"
colnames(soil_iso_data3)[6] = "total_n_ug"
colnames(soil_iso_data3)[2] = "d13c"
colnames(soil_iso_data3)[5] = "d15n"
colnames(soil_iso_data3)[12] = "sample_weight_mg"
soil_iso_data4 <- soil_iso_data3 %>% 
  subset(!is.na(lake))
soil_iso_data4 <- soil_iso_data4 %>% 
  subset(!is.na(pct_OM)) + 
  substr

```

Calculate percent C and percent N
```{r}
# Calculate percent N
soil_iso_data4 <- soil_iso_data4 %>%
  mutate(pct_n = total_n_ug / (as.numeric(sample_weight_mg)*1000) * 100)
View(soil_iso_data4)

# Calculate percent C
soil_iso_data4 <- soil_iso_data4 %>%
  mutate(pct_c = total_c_ug / (as.numeric(sample_weight_mg)*1000) * 100)
View(soil_iso_data4)

```

Graphs of percent C and percent N in soils based on elevation and fish
```{r}
# Make scatterplot of percent C as a function of elevation and fish
soil_iso_data4 %>% 
  ggplot(aes(x = elevation_m, y = pct_c, color = fish_status)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm <- lm(pct_c ~ elevation_m + fish_status, data = soil_iso_data4)
summary(res.lm) # not significant

# Make scatterplot of percent N as a function of elevation and fish  
soil_iso_data4 %>% 
  ggplot(aes(x = elevation_m, y = pct_n, color = fish_status)) +
  geom_jitter() +
  geom_smooth(method="lm") # fish presence increases percent N, percent N increases with elevation

# Test significance
res.lm2 <- lm(pct_n ~ elevation_m + fish_status, data = soil_iso_data4)
summary(res.lm2) # fish presence and elevation (but not interaction) significantly increase percent N
```

Graphs of d13C and d15N by elevation for soils
```{r}
# Make scatterplot of d13C as a function of elevation and fish
soil_iso_data4 %>% 
  ggplot(aes(x = elevation_m, y = d13c, color = fish_status)) +
  geom_jitter() +
  geom_smooth(method="lm")

# Test significance
res.lm3 <- lm(d13c ~ elevation_m * fish_status, data = soil_iso_data4) # elevation and fish status both (and interaction) significantly increase d13c, elevation more than fish status
summary(res.lm3)

# Make scatterplot of d15N as a function of distance and fish
soil_iso_data4 %>% 
  ggplot(aes(x = elevation_m, y = d15n, color = fish_status)) +
  geom_jitter() +
  geom_smooth(method="lm") # fish and elevation both seem to increase d15n

# Test significance
res.lm4 <- lm(d15n ~ elevation_m + fish_status, data = soil_iso_data4)
summary(res.lm4) # fish and elevation (but not interaction) significantly increase d15n in soils

```

Graphs of percent N by percent OM
```{r}
# Make scatterplot of percent N as a function of percent OM
soil_iso_data4 %>% 
  ggplot(aes(x = pct_OM, y = pct_n, color = fish_status)) +
  geom_jitter() +
  geom_smooth(method="lm")

res.lm5 <- lm(pct_n ~ pct_OM, data = soil_iso_data4) # pct_OM significantly increases pct_n in soils
summary(res.lm5)
```

Interactions between N as a function of OM, elevation, fish status
```{r}

res.lm6 <- lm(sqrt(pct_n) ~ pct_OM * elevation_m * fish_status, data = soil_iso_data4)
summary(res.lm6)
qqPlot(res.lm6)
ggpredict(res.lm6, terms = c("fish_status", "elevation_m"))

res.lm7 <- lm(sqrt(pct_n) ~ pct_OM * (elevation_m + fish_status), data = soil_iso_data4) 
summary(res.lm7)
qqPlot(res.lm7)

res.lm8 <- lm(sqrt(pct_n) ~ pct_OM + elevation_m + fish_status, data = soil_iso_data4) 
summary(res.lm8)
qqPlot(res.lm8)

res.lm9 <- lm(sqrt(pct_n) ~ pct_OM * elevation_m, data = soil_iso_data4) 
summary(res.lm9)
qqPlot(res.lm9)

res.lm10 <- lm(sqrt(pct_n) ~ pct_OM + elevation_m, data = soil_iso_data4) 
summary(res.lm10)
qqPlot(res.lm10)

AIC(res.lm6, res.lm7, res.lm8, res.lm9, res.lm10)

```
